Temporal graph realization from fastest paths

### 2024-01-17 SAND2024/SAND24-longversion.tex
changing how we guess fastest temporal paths (added observation that all segments are determined if they are of length at least 4)

Formally, let $G=(V,E)$ and $\Delta\in \mathbb{N}$, and let $\lambda: E \rightarrow \{1,2,\ldots,\Delta\}$ be an edge-labeling function that assigns to every edge of $G$ exactly one of the labels from $\{1,\ldots,\Delta\}$. 
Then we denote by $(G,\lambda,\Delta)$ the \emph{$\Delta$-periodic temporal graph} $(G,L)$, where for every edge $e\in E$ we have $L(e)=\{i\Delta + x : i\geq 0, x\in \lambda(e)\}$. 
In this case we call $\lambda$ a \emph{$\Delta$-periodic labeling} of~$G$; see \cref{fig:periodic-example} for an illustration. 
When it is clear from the context, we drop $\Delta$ form the notation and 
we denote the ($\Delta$-periodic) temporal graph by $(G,\lambda)$.
Given a duration matrix~$D$, it is easy to observe that, similarly to the static case, if $D_{i,j}=1$ then $v_i$ and $v_j$ must be connected by an edge. We call the graph defined by these edges the \emph{underlying graph} of $D$.

@@

Recall from \cref{sec:treealgo} that the main reason, for which \deltaExact\ is straightforward to solve on trees, is twofold: 
\begin{itemize}
    \item between any pair of vertices $v_i$ and $v_j$ in the tree $T$, there is a \emph{unique} path $P$ in $T$ from $v_i$ to $v_j$, and 
    \item in any periodic temporal graph $(T,\lambda,\Delta)$ and any fastest temporal path $P=((e_1,t_1),\ldots,(e_i,t_i),\ldots,(e_j,t_j),\ldots,(e_\ell,t_\ell))$ from $v_1$ to $v_{\ell}$ we have that the sub-path $P'=((e_i,t_i),\ldots,(e_j,t_j))$ is also a fastest temporal path from $v_i$ to $v_j$.
\end{itemize}
However, these two nice properties do not hold when the underlying graph is not a tree. For example, in~\cref{fig:ftpExample}, the fastest temporal path from $u$ to $v$ is $P_{u,v}$ (depicted in blue) goes through~$w$, however the sub-path of $P_{u,v}$ that stops at $w$ is not the fastest temporal path from $u$ to $w$. The fastest temporal path from $u$ to $w$ consists only of the single edge $uw$ (with label $9$ and duration $1$, depicted in red).
%, while the fastest temporal path from $u$ to $v$ contains the four other edges (with labels 1, 3, 5, 7 and duration 7).

@@

From the construction of $Z$ (\ie by exhaustively removing vertices of degree one from~$G$), it follows that $G[Z]$ (the graph induced in $G$ by $Z$) consists of disjoint trees. 
Each of these trees has a unique neighbor $v$ in $G'$. 
Denote by $T_v$ the tree obtained by considering such a vertex $v$ and all the trees from $G[Z]$ that are incident to $v$ in $G$. 
We then refer to $v$ as the \emph{clip vertex} of the tree $T_v$.
In the case $v$ is a vertex of interest we define also the set $Z^*_v$ of \emph{representative vertices} of $T_v$, as follows.

\noindent
We first create an empty set $C_w$ for every vertex $w$ that is a neighbor of $u$ in $G'$.
We then iterate through every vertex $r$ that is in the first layer of the tree $T_v$ (\ie vertex that is a direct child from the root $v$), we check the matrix $D$ and find the vertex $w \in N_{G'}(v)$ that is on the shortest duration from $r$.
In other words, for a $r \in N_{T_v}(v)$ we find $w \in N_{G'}(v)$ such that $D_{r,w} \leq D_{r,w'}$ for all $w' \in  N_{G'}(v)$.
%
We add vertex $r$ to $C_w$.
In the case when there exist another $w'' \in  N_{G'}(v) $ for which $D_{r,w''} = D_{r,w}$, we add $r$ also to the set $C_{w''}$. In fact, in this case $C_{w''} = C_w$.
At the end we create $|N_{G'}(v)|$ sets $C_w$, containing all vertices of the first layer of $T_v$.
We call these sets \emph{equivalence classes} of the tree $T_v$.
Now, from each equivalence class $C_w$ we chose an arbitrary vertex $r_w$ and put it into the set $Z^*_v$.
%
We repeat the above procedure for all trees $T_u$ with the clip vertex $u$ from $U$, and define $Z^*$ as 
$$Z^* = \bigcup\limits_{u \in U} Z^*_u.$$
%
Since $|U| \in O(k)$ and for each $u$ it holds $|N(u)| \in O(k)$, we get that $|Z^*| \in O(k^2)$. 
%For every vertex in the first layer of the tree $T_u$, where $u \in U$, we check the corresponding values in the matrix $D$ exactly once. Therefore, $Z^*$ is determined in linear time.

Finally, the set of \emph{important vertices} is defined as the set $U \cup U^{\ast} \cup Z^{\ast}$.
For an illustration see \cref{fig:labelingVertices}. Note that determining sets $U, U^*$ and $Z^*$ takes linear time.

@@

%Let us remember the following basic property of the problem \deltaExact.
Recall that a labeling $\lambda$ of $G$ satisfies $D$ if the duration of a fastest temporal path from vertex $v_i$ to $v_j$ equals $D_{v_i, v_j}$.
%, and no other temporal path from $v_i$ to $v_j$ is faster.
In order to find a labeling that satisfies this property we split our analysis in nine cases.
We consider the fastest temporal paths 
where the starting vertex is in one of the sets $U,V(G') \setminus U, Z$,
and similarly the destination vertex is in one of the sets $U,V(G') \setminus U, Z$.
In each of these cases, we guess the underlying path $P$ that at least one fastest temporal path from the vertex $v_i$ to $v_j$ follows, 
which results in one equality constraint for the labels on the path $P$. 
For all other temporal paths from $v_i$ to $v_j$ we know that they cannot be faster, so we introduce inequality constraints for them.
This results in producing $f(k) \cdot |D| ^ {O(1)}$ constraints. 
Note that we have to do this while keeping the total number of variables upper-bounded by some function in $k$.
%We have to guarantee only that the number of guesses of fastest temporal paths is exponential in $k$ and not in $n$.

For an easier understanding and analysis of the algorithm, we give the following definition.
\begin{definition}\label{def:segments}
    Let $U \subseteq V(G')$ be a set of vertices of interest and let $u,v \in U$.
    A path $P = (u=v_1,v_2, \dots, v_p = v)$  in graph $G'$, where all inner vertices are not in $U$, \ie $v_i \notin U$ for all $i \in \{ 2, 3, \dots, p-1\}$,
    is called a \emph{segment} from $u$ to $v$. We denote it as $S_{u,v}$.
\end{definition}
%TODO: We can have two segments among a pair of vertices in U !!!
%Among a fixed pair of vertices $u,v \in U$ there can be at most two different segments. If there are two, then one has to be of length $1$, just a single edge, and the second one of length at least $2$.
Note from \cref{def:segments} that $S_{u,v} \neq S_{v,u}$ since we consider paths to be directed. It's also worth emphasizing that $S_{v,u}$ is essentially the reversed path of $S_{u,v}$.
Furthermore, it's important to observe that a temporal path in $G'$ between two vertices of interest is either a segment or consists of a sequence of segments.
Moreover, any inner vertex $v_i$ in the segment $S_{u,v}$ ($v_i \in S_{u,v} \setminus \{u,v\}$) is part of precisely two segments: $S_{u,v}$ and $S_{v,u}$.
Given that we have at most $4k$ interesting vertices in $G'$, we can deduce the following crucial result.
\begin{corollary}\label{obs:FPT-k2segments}
    There are at most $O(k^2)$ segments in $G'$.
\end{corollary}
%TODO: one reviewer asked how we get this number? (number of paths)
%For a pair $u,v$ of vertices in $U$, there are $O(k!)=O(k^k)$ possible paths in $G'$ between them.

\subsubsection{Guessing necessary structures \label{sec:FPT-guessing}}
Once the sets $U, U^*$ and $Z^*$ are determined, we are ready to start guessing  the necessary structures.
Note that whenever we say that we guess the fastest temporal path between two vertices, we mean that we guess the underlying path of a representative fastest temporal path between those two vertices.
To describe the guesses, we introduce the following notation. Let $u,v,x$ be three vertices in $G'$. We write $u \leadsto x \rightarrow v$ to denote a temporal path from $u$ to $v$ that passes through $x$, and then goes directly to $v$ (via one edge). If there is an edge (\ie a unique fastest path) between two vertices, we denote it by $\rightarrow$,
if the fastest path between two vertices is not uniquely determined, we denote it by $\leadsto$.

%For every pair of important vertices $u,v \in U \cup U^{\ast} \cup Z^{\ast}$, we guess the sequence of edges in the fastest temporal path from $u$ to $v$. Since $U \cup U^{\ast} \cup Z^{\ast}\in O(k^2)$ and there are $k^{O(k)}$ possibilities for a sequence of edges between a fixed vertex pair, we have $k^{O(k^5)}$ overall possible guesses.
We start by guessing the following.
\begin{enumerate}[G-1.]
	\setcounter{enumi}{\value{guesscounter}}
    \item \label{FPT-guessFTPamongU}
    The fastest temporal paths between all pairs of vertices of $U$.
    For a pair $u,v$ of vertices in $U$, there are $k^{O(k)}$ possible paths in $G'$ between them. 
    Therefore, we have to try all $k^{O(k)}$ possible paths, where at least one of them will be a fastest temporal path from $u$ to $v$, respecting the values from $D$.
    %If this does not happen we conclude that $D$ is a \textsc{No} instance of \deltaExact.
    Repeating this procedure for all pairs of vertices $u,v \in U$ we get $k^{O(k^3)}$ different variations of the fastest temporal paths between all pairs of vertices in $U$.
    \item \label{FPT-guessFTPamongZstar}
    The fastest temporal paths between all pairs of vertices in $Z^*$, 
    which by similar arguing as for vertices in $U$, gives us $k^{O(k^5)}$ guesses.
    \item \label{FPT-guessFTPamongUstar}
    The fastest temporal paths between all pairs of vertices in $U^*$.
    This gives us $k^{O(k^5)}$ guesses.
    \item \label{FPT-guessFTPamongUandUstar}
    The fastest temporal paths from vertices of $U$ to vertices in $U^*$,
    and vice versa, the fastest temporal paths from vertices in $U^*$ to vertices in $U$.
    This gives us $k^{O(k^4)}$ guesses.
    \item \label{FPT-guessFTPamongUandZstar}
    The fastest temporal paths from vertices of $U$ to vertices in $Z^*$,
    and vice versa.
    This gives us $k^{O(k^4)}$ guesses.
    \item \label{FPT-guessFTPamongUstarandZstar}
    The fastest temporal paths from vertices of $U^*$ to vertices in $Z^*$,
    and vice versa.
    This gives us $k^{O(k^5)}$ guesses.
	\setcounter{guesscounter}{\value{enumi}}
\end{enumerate}

@@

Therefore we introduce additional guesses that provide us with sufficient information to determine all fastest paths.

@@

We now present some properties that involve the vertices from $Z$, that form the trees in $G[Z]$.
\begin{lemma}\label{lemma:ftp-tree-neighbor}
Let $v \in V(G')$ be a clip vertex of the tree $T_v$ in $G[Z \cup \{v\}]$, and let $r_i \in N_{T_v}(v)$ be an arbitrary vertex of $T_v$ that is a neighbor (child) of $v$. 
Among all neighbors of $v$ in $G'$ let $w$ be the one that is on the smallest duration away from $r_i$ with respect to the values of $D$. 
In other words, $w \in N_{G'}(v)$ such that $D_{r_i,w} \leq D_{r_i,w'}$ for all $w' \in N_{G'}(v)$.
Then, the path $P^*=(r_i,v,w)$ represents the fastest temporal path from $r_i$ to $w$.
%
Moreover, all labels of the tree $T_v$ are determined with respect to the label $\lambda(vw)$.
\end{lemma}
\begin{proof}
Suppose for contradiction that there exists a path $P^*_2$ from $r_i$ to $w$ that is faster than $P^*=(r_i,v,w)$. 
By the structure of $G$, it follows that $P^*_2$ passes through the clip vertex $v$ of $T_v$ (as this is the only neighbor of $r_i$ in $G'$), continues through a vertex $w' \in N_{G'}(v) \setminus \{w \}$, 
and through some other vertices $u_1, u_2, \dots, u_j$ in $G$ ($j \geq 0$) before finishing in $w$.
Therefore, $P^*_2= (r_i,v,w',u_1,u_2,\dots,u_j,w)$, and $d(P^*_2,\lambda) \leq d(P^*,\lambda)$.
Now, since $D_{r_i,w} \leq D_{r_i,w'}$ for all $ w' \in N_{G'}(v)$ the first part of path $P^*_2$ from $r_i$ to $w'$ takes at least $D_{r_i,w}$ time.
Since $w \neq w'$, we need at least one more time-step (one more edge) to traverse from $w'$ to reach $w$. 
Therefore, $d(P^*_2, \lambda) \geq D_{r_i,w} + 1$ which implies that $P^*$ is a faster temporal path than $P^*_2$.

For the second part, knowing that the duration of $P^*$ is $D_{r_i,w}$ we can determine the label of the edge $r_i v$ with respect to the label $\lambda(vw)$ (see \cref{obs:FirstLabelAndDuration}).
Furthermore, using the algorithm for trees (see \cref{thm:deltaExact-PolyTimeTrees}), we can now determine labels on all edges of $T_v$ with respect to the same label $\lambda(v w)$.
\end{proof}

\begin{lemma} \label{lemma:treeLabels-notInteresting}
    Let $z \in V(G')$ be a clip vertex of the tree $T_i$ in $G[Z \cup \{z\}]$, where $z \notin U$. 
    Let $v_1$ and $v_2$ be the two neighbors of $z$ in $G'$.
    Then the labels of the tree $T_i$ can be determined with respect to $\lambda(v_1 z)$ and $\lambda(z v_2)$.
\end{lemma}
%
\begin{proof}
    First observe that since $z$ is not a vertex of interest it must be a part of some segment $S_{u,w}$, where $u,v \in U$ and $z \neq u \neq v$. Therefore, $z$ is of degree $2$ in $G'$.
    Let $r \in V(T_i)$ be a neighbor of $z$.
    We observe the values $D_{r,v_1}, D_{r,v_2}$ and distinguish the following cases.

    First, $D_{r,v_1} = D_{r,v_2}$
    Then, using \cref{lemma:ftp-tree-neighbor} we get that the fastest paths from $r$ to $v_1$ and from $r$ to $v_2$ are of length two.
    Moreover, we know that these two paths consist of the edge $rz$ and one of the edges $zv_1,zv_2$.
    This allows us to determine the label of the edge $rz$ (and consequently all other edges of $T_i$) with respect to $\lambda(zv_1)$ (see \cref{obs:FirstLabelAndDuration}).

    Second, $D_{r,v_1} \neq D_{r,v_2}$. 
    Let us denote with $t_1 = \lambda(z v_1), t_2 = \lambda(z v_2)$ and $t_3 = \lambda(rz)$.
    W.l.o.g. suppose $D_{r,v_1} > D_{r,v_2}$, which implies that $t_1 > t_2$ (the other case is analogous), and suppose that $t_1, t_2 > t_3$.
    Now we claim that $D_{v_1,r} < D_{v_2,r}$.
    Suppose for the contradiction that this is not true. Then $D_{v_2,r} < D_{v_1,r} \leq (\Delta + t_3 - t_1)$. This implies that the fastest temporal path from $v_2$ to $r$ cannot use the path $(v_1,z,r)$, and is therefore of form $(v_2,z,r)$.
    By the definition, the duration of this path is then $D_{v_2,r} = \Delta + t_3 - t_2 + 1$.
    But since $t_1 > t_2$ it follows that $(\Delta + t_3 - t_2) + 1 > (\Delta + t_3 - t_1) + 1$ and by the definition $D_{v_1,r} \leq > (\Delta + t_3 - t_1) + 1$. 
    Therefore we get $D_{v_2,r} > D_{v_1,r} $, a contradiction.
    All together, from $D_{r,v_1} > D_{r,v_2}$ we determine the label of edge $rz$ with respect to $\lambda(z v_2)$,
    and from $D_{v_1,r} < D_{v_2,r}$ we determine the label of edge $rz$ with respect to $\lambda(z v_1)$.
    %
\end{proof}

%%%% TODO NKL some description of equivalence classes
Let us now show the following important property of equivalence classes.
\begin{lemma}\label{lemma:treeEquivClass-intVertices}
    Let $u \in V(G')$ be a clip vertex of the tree $T_i$ in $G[Z \cup \{u\}]$, where $u \in U$,
    and let $r_i,r_2 \in V(T_i)$ be in the same equivalence class of the tree $T_i$.
    Then, the fastest temporal paths from $r_1$ and $r_2$ to any other vertex in $G$ coincide on the vertices of $V(G) \setminus V(T_i)$.
    Similarly, it holds for the fastest temporal paths from any other vertex in $G$ to $r_1$ and $r_2$.
\end{lemma}
%
\begin{proof}
    Let $y \neq u$ be a vertex in $V(G')$.
    %
    Denote with $P_1$ the underlying path of the fastest temporal path from $r_1$ to $y$, which consists of the edge $r_1 u$ and the path $P$ from $u$ to $y$.
    Similarly, let $Q_2$ be the underlying path of the fastest temporal path from $r_2$ to $y$, consisting of the edge $r_2u$ together with the path $Q$ from $u$ to $y$.
    Define $P_2$ as the second path from $r_2$ to $y$ that first uses the edge $r_1u$ and then the path $P$.
    Similarly, $Q_1$ represents the second path from $r_1$ to $y$ that first uses the edge $r_2u$ and then the path $Q$.
    Our objective is to demonstrate that either $P = Q$ or that $d(P_1, \lambda) = d(Q_1, \lambda)$ (and $d(P_2, \lambda) = d(Q_2, \lambda)$).
    This implies that $r_1$ and $r_2$ use temporal paths that coincide on the vertices of $V(G) \setminus V(T_i)$ to reach $y$.

>>>>

Formally, let $G=(V,E)$ and $\Delta\in \mathbb{N}$, and let $\lambda: E \rightarrow \{1,2,\ldots,\Delta\}$ be an edge-labeling function that assigns to every edge of $G$ exactly one of the labels from $\{1,\ldots,\Delta\}$. 
Then we denote by $(G,\lambda,\Delta)$ the \emph{$\Delta$-periodic temporal graph} $(G,L)$, where for every edge $e\in E$ we have $L(e)=\{i\Delta + x : i\geq 0, x\in \lambda(e)\}$. 
In this case we call $\lambda$ a \emph{$\Delta$-periodic labeling} of~$G$; see \cref{fig:periodic-example} for an illustration. 
When it is clear from the context, we drop $\Delta$ from the notation and 
we denote the ($\Delta$-periodic) temporal graph by $(G,\lambda)$.
Given a duration matrix~$D$, it is easy to observe that, similarly to the static case, if $D_{i,j}=1$ then $v_i$ and $v_j$ must be connected by an edge. We call the graph defined by these edges the \emph{underlying graph} of $D$.

@@

Recall from \cref{sec:treealgo} that the main reason, for which \deltaExact\ is straightforward to solve on trees, is twofold: 
\begin{itemize}
    \item between any pair of vertices $v_i$ and $v_j$ in the tree $T$, there is a \emph{unique} path $P$ in $T$ from $v_i$ to $v_j$, and 
    \item in any periodic temporal graph $(T,\lambda,\Delta)$ and any fastest temporal path $P=((e_1,t_1),\ldots,(e_i,t_i),\ldots,(e_j,t_j),\ldots,(e_{\ell-1},t_{\ell-1}))$ from $v_1$ to $v_{\ell}$ we have that the sub-path $P'=((e_i,t_i),\ldots,(e_{j-1},t_{j-1}))$ is also a fastest temporal path from $v_i$ to $v_j$.
\end{itemize}
However, these two nice properties do not hold when the underlying graph is not a tree. For example, in~\cref{fig:ftpExample}, the fastest temporal path from $u$ to $v$ is $P_{u,v}$ (depicted in blue) goes through~$w$, however the sub-path of $P_{u,v}$ that stops at $w$ is not the fastest temporal path from $u$ to $w$. The fastest temporal path from $u$ to $w$ consists only of the single edge $uw$ (with label $9$ and duration $1$, depicted in red).
%, while the fastest temporal path from $u$ to $v$ contains the four other edges (with labels 1, 3, 5, 7 and duration 7).

@@

From the construction of $Z$ (\ie by exhaustively removing vertices of degree one from~$G$), it follows that $G[Z]$ (the graph induced in $G$ by $Z$) consists of disjoint trees. 
Each of these trees has a unique neighbor $v$ in $G'$. 
Denote by $T_v$ the tree obtained by considering such a vertex $v$ and all the trees from $G[Z]$ that are incident to $v$ in $G$. 
We then refer to $v$ as the \emph{clip vertex} of the tree $T_v$.
In the case where $v$ is a vertex of interest we define also the set $Z^*_v$ of \emph{representative vertices} of $T_v$, as follows.

\noindent
We first create an empty set $C_w$ for every vertex $w$ that is a neighbor of $u$ in $G'$.
We then iterate through every vertex $r$ that is in the first layer of the tree $T_v$ (\ie vertex that is a child of the root $v$ in the tree $T_v$), check the matrix $D$ and find the vertex $w \in N_{G'}(v)$ that is on the smallest duration from $r$.
In other words, for an $r \in N_{T_v}(v)$ we find $w \in N_{G'}(v)$ such that $D_{r,w} \leq D_{r,w'}$ for all $w' \in  N_{G'}(v)$.
%
We add vertex $r$ to $C_w$.
In the case when there exists also another vertex $w' \in  N_{G'}(v) $ for which $D_{r,w'} = D_{r,w}$, we add $r$ also to the set $C_{w'}$. In fact, in this case $C_{w'} = C_w$.
At the end we create $|N_{G'}(v)|=O(k)$ sets $C_w$, whose union contains all children of $v$ in $T_v$. 
For every two sets $C_w$ and $C_{w'}$, where $w,w'\in N_{G'}(v)$, we have that either $C_w = C_{w'}$, or $C_w \cap C_{w'} = \emptyset$. 
We call each of these sets $\{C_w : w \in N_{G'}(v)\}$ an \emph{equivalence class} of the tree $T_v$. 
Now, from each equivalence class $C_w$ we choose an arbitrary vertex $r_w$ and put it into the set $Z^*_v$.
%
We repeat the above procedure for all trees $T_u$ with the clip vertex $u$ from $U$, and define $Z^*$ as 
\begin{equation}
    Z^* = \bigcup\limits_{v \in U} Z^*_v.
\end{equation}
%
Since $|U| \in O(k)$ and for each $u \in U$ it holds $|N_{G'}(u)| = O(k)$, we get that $|Z^*| \in O(k^2)$. 
%For every vertex in the first layer of the tree $T_u$, where $u \in U$, we check the corresponding values in the matrix $D$ exactly once. Therefore, $Z^*$ is determined in linear time. 
Finally, the set of \emph{important vertices} is defined as the set $U \cup U^{\ast} \cup Z^{\ast}$.
For an illustration see \cref{fig:labelingVertices}. Note that determining sets $U, U^*$ and $Z^*$ takes linear time.

@@

%Let us remember the following basic property of the problem \deltaExact.
Recall that a labeling $\lambda$ of $G$ satisfies $D$ if the duration of a fastest temporal path from each vertex $v_i$ to each other vertex $v_j$ equals $D_{v_i, v_j}$.
%, and no other temporal path from $v_i$ to $v_j$ is faster.
In order to find a labeling that satisfies this property we split our analysis in nine cases.
We consider the fastest temporal paths 
where the starting vertex is in one of the sets $U,V(G') \setminus U, Z$,
and similarly the destination vertex is in one of the sets $U,V(G') \setminus U, Z$.
In each of these cases, we guess the underlying path $P$ that at least one fastest temporal path from the vertex $v_i$ to $v_j$ follows, 
which results in one equality constraint for the labels on the path $P$. 
For all other temporal paths from $v_i$ to $v_j$ we know that they cannot be faster, so we introduce inequality constraints for them.
This results in producing $f(k) \cdot |D| ^ {O(1)}$ constraints. 
Note that we have to do this while keeping the total number of variables upper-bounded by some function in $k$.
%We have to guarantee only that the number of guesses of fastest temporal paths is exponential in $k$ and not in $n$.

For an easier understanding and analysis of the algorithm, we give the following definition.
\begin{definition}\label{def:segments}
    Let $U \subseteq V(G')$ be a set of vertices of interest and let $u,v \in U$.
    A path $P = (u=v_1,v_2, \dots, v_p = v)$  with at least two edges in graph $G'$, where all inner vertices are not in $U$, \ie $v_i \notin U$ for all $i \in \{ 2, 3, \dots, p-1\}$,
    is called a \emph{segment} from $u$ to $v$, which we denote as $S_{u,v}$.
\end{definition}
%TODO: We can have two segments among a pair of vertices in U !!!
%Among a fixed pair of vertices $u,v \in U$ there can be at most two different segments. If there are two, then one has to be of length $1$, just a single edge, and the second one of length at least $2$.
Note from \cref{def:segments} that $S_{u,v} \neq S_{v,u}$ since we consider paths to be directed. It is also worth emphasizing that $S_{v,u}$ is essentially the reverse path of $S_{u,v}$.
Furthermore, it's important to observe that a temporal path in $G'$ between two vertices of interest is either a segment or consists of a sequence of segments.
Moreover, any inner vertex $v_i$ in the segment $S_{u,v}$ ($v_i \in S_{u,v} \setminus \{u,v\}$) is part of precisely two segments: $S_{u,v}$ and $S_{v,u}$.
Given that we have at most $4k$ interesting vertices in $G'$, we can deduce the following crucial result.
\begin{corollary}\label{obs:FPT-k2segments}
    There are at most $O(k^2)$ segments in $G'$.
\end{corollary}
%TODO: one reviewer asked how we get this number? (number of paths)
%For a pair $u,v$ of vertices in $U$, there are $O(k!)=O(k^k)$ possible paths in $G'$ between them.

\subsubsection{Guessing necessary structures \label{sec:FPT-guessing}}
Once the sets $U, U^*$ and $Z^*$ are determined, we are ready to start guessing  the necessary structures.
Note that whenever we say that we guess the fastest temporal path between two vertices, we mean that we guess the underlying path of a representative fastest temporal path between those two vertices.
To describe the guesses, we introduce the following notation. Let $u,v,x$ be three vertices in $G'$. We write $u \leadsto x \rightarrow v$ to denote a temporal path from $u$ to $v$ that passes through $x$, and then goes directly to $v$ (via one edge or a unique path in $G'$). 
In other words, if the fastest path between two vertices is not uniquely determined we denote it by $\leadsto$, while if it is unique we denote it by $\rightarrow$.
%
%For every pair of important vertices $u,v \in U \cup U^{\ast} \cup Z^{\ast}$, we guess the sequence of edges in the fastest temporal path from $u$ to $v$. Since $U \cup U^{\ast} \cup Z^{\ast}\in O(k^2)$ and there are $k^{O(k)}$ possibilities for a sequence of edges between a fixed vertex pair, we have $k^{O(k^5)}$ overall possible guesses.
We guess the following paths.
\begin{enumerate}[G-1.]
	\setcounter{enumi}{\value{guesscounter}}
    \item \label{FPT-guessFTPamongU}
    The fastest temporal paths between all pairs of vertices of $U$.
    For a pair $u,v$ of vertices in $U$, there are $k^{O(k)}$ possible paths in $G'$ between them. 
    Therefore, we have to try all $k^{O(k)}$ possible paths, where at least one of them will be a fastest temporal path from $u$ to $v$, respecting the values from $D$.
    %If this does not happen we conclude that $D$ is a \textsc{No} instance of \deltaExact.
    Repeating this procedure for all pairs of vertices $u,v \in U$ we get $k^{O(k^3)}$ different variations of the fastest temporal paths between all pairs of vertices in $U$.
    \item \label{FPT-guessFTPamongZstar}
    The fastest temporal paths between all pairs of vertices in $Z^*$, 
    which by similar arguing as for vertices in $U$, gives us $k^{O(k^5)}$ guesses.
    \item \label{FPT-guessFTPamongUstar}
    The fastest temporal paths between all pairs of vertices in $U^*$.
    This gives us $k^{O(k^5)}$ guesses.
    \item \label{FPT-guessFTPamongUandUstar}
    The fastest temporal paths from vertices of $U$ to vertices in $U^*$,
    and vice versa, the fastest temporal paths from vertices in $U^*$ to vertices in $U$.
    This gives us $k^{O(k^4)}$ guesses.
    \item \label{FPT-guessFTPamongUandZstar}
    The fastest temporal paths from vertices of $U$ to vertices in $Z^*$,
    and vice versa.
    This gives us $k^{O(k^4)}$ guesses.
    \item \label{FPT-guessFTPamongUstarandZstar}
    The fastest temporal paths from vertices of $U^*$ to vertices in $Z^*$,
    and vice versa.
    This gives us $k^{O(k^5)}$ guesses.
	\setcounter{guesscounter}{\value{enumi}}
\end{enumerate}

@@

Therefore, we introduce additional guesses that provide us with sufficient information to determine all fastest paths.

@@

We now present some properties involving the vertices from $Z$, that form the trees in~$G[Z]$.
\begin{lemma}\label{lemma:ftp-tree-neighbor}
Let $v \in V(G')$ be a clip vertex of the tree $T_v$ in $G[Z \cup \{v\}]$, and let $z \in N_{T_v}(v)$ be an arbitrary child of $v$ in $T_v$. 
Among all neighbors of $v$ in $G'$, let $w$ be the one that is on the smallest duration away from $z$ with respect to the values of $D$. 
In other words, $w \in N_{G'}(v)$ such that $D_{z,w} \leq D_{z,w'}$ for all $w' \in N_{G'}(v)$.
Then, the path $P^*=(z,v,w)$ represents the unique fastest temporal path from $z$ to $w$.
%
Moreover, we can determine all labels of the tree $T_v$ with respect to the label $\lambda(vw)$.
\end{lemma}
\begin{proof}
Suppose for contradiction that there exists a path $P^{**} \neq P^*$ from $z$ to $w$ such that $d(P^{**},\lambda) \leq d(P^*,\lambda)$. 
By the structure of $G$, it follows that $P^{**}$ passes through the clip vertex $v$ of $T_v$ (as this is the only neighbor of $z$ in $G'$), continues through a vertex $w' \in N_{G'}(v) \setminus \{w \}$, 
and through some other vertices $u_1, u_2, \dots, u_j$ in $G$ ($j \geq 0$) before finishing in $w$.
Therefore, $P^{**}= (z,v,w',u_1,u_2,\dots,u_j,w)$.
Now, since $D_{z,w} \leq D_{z,w'}$ by assumption, the first part of $P^{**}$ from $z$ to $w'$ takes at least $D_{z,w'}$ time, and thus it takes at least $D_{z,w}$ time.
Since $w \neq w'$, we need at least one more time-step (one more edge) to traverse from $w'$ to reach $w$. 
Therefore, $d(P^{**}, \lambda) \geq D_{z,w} + 1$ 
which implies that $P^{**}$ is not the fastest temporal path from $z$ to $w$. Therefore, the only fastest temporal path from $z$ to $w$ is $P^*=(z,v,w)$.

For the second part, knowing that the duration of $P^*$ is $D_{z,w}$, we can determine the label of the edge $z v$ with respect to the label $\lambda(vw)$ (see \cref{obs:FirstLabelAndDuration}).
Furthermore, using the algorithm for trees (see \cref{thm:deltaExact-PolyTimeTrees}), we can now determine all the labels on the edges of $T_v$ with respect to the same label $\lambda(v w)$.
\end{proof}

\begin{lemma} \label{lemma:treeLabels-notInteresting}
    Let $x \in V(G')$ be a clip vertex of the tree $T_x$ in $G[Z \cup \{x\}]$, where $x \notin U$. 
    Let $v_1$ and $v_2$ be the two neighbors of $x$ in $G'$.
    Then the labels of the tree $T_x$ can be determined with respect to $\lambda(v_1 x)$ and $\lambda(x v_2)$.
\end{lemma}
%
\begin{proof}
    First observe that since $x$ is not a vertex of interest it must be a part of some segment $S_{u,w}$, where $u,v \in U$ and $x \neq u \neq v$. Therefore, $x$ is of degree $2$ in $G'$.
    Let $z \in V(T_x)$ be a child of $x$ in $T_x$, \ie a vertex in the first layer of the tree $T_x$.
    We observe the values $D_{z,v_1}, D_{z,v_2}$ and distinguish the following cases.

    First, $D_{z,v_1} = D_{z,v_2}$
    Then, using \cref{lemma:ftp-tree-neighbor} we conclude that the fastest temporal paths from $z$ to $v_1$ and from $z$ to $v_2$ are of length two.
    We know that these two paths consist of the edge $zx$ and $xv_1,xv_2$, respectively.
    This allows us to determine the label of the edge $zx$ (and consequently all other edges of $T_x$) with respect to $\lambda(xv_1)$ and $\lambda(xv_2)$.

    Second, $D_{z,v_1} \neq D_{z,v_2}$. 
    Let us denote with $t_1 = \lambda(x v_1), t_2 = \lambda(x v_2)$ and $t_3 = \lambda(zx)$.
    W.l.o.g. suppose that $\min \{t_1,t_2,t_3\} = t_3$, and that $D_{z,v_1} > D_{z,v_2}$ (the other case is analogous). It follows that $t_1 > t_2$.
    We want to now prove that the inequality $D_{v_1,z} < D_{v_2,z}$ holds.
    Suppose for the contradiction that the inequality is false. Then $D_{v_2,z} < D_{v_1,z} \leq (\Delta + t_3 - t_1)$. This implies that the fastest temporal path from $v_2$ to $z$ cannot use the path $(v_1,x,z)$, and is therefore of form $(v_2,x,z)$.
    By the definition, the duration of this path is $D_{v_2,z} = \Delta + t_3 - t_2 + 1$.
    But since $t_1 > t_2$ it follows that $(\Delta + t_3 - t_2) + 1 > (\Delta + t_3 - t_1) + 1$.  We also know that $D_{v_1,z} \leq  (\Delta + t_3 - t_1) + 1$. 
    This implies that $D_{v_2,z} > D_{v_1,z} $, a contradiction.\\
    Knowing $D_{z,v_1} > D_{z,v_2}$ we can determine the label of edge $zx$ (and consequently all other edges of $T_x$) with respect to $\lambda(x v_2)$,
    and similarly knowing $D_{v_1,z} < D_{v_2,z}$ we determine the label of edge $zx$ (and all other edges of $T_x$) with respect to $\lambda(x v_1)$.
    %
\end{proof}
